Theories of deliberated choice

### 2023-06-21 simple.tex
Towards proving equivalence with edge case

We now say that $C$ is \emph{falsifiable} iff $\forall {\allleadstog} \in P^{I × \allhist}: ¬C({\allleadstog}) \implies \exists O \subseteq {\allleadstog} \suchthat O \implies ¬C$, meaning that
$\forall {\allleadstog} \notin C: \exists O \subseteq {\allleadstog} \suchthat S_O \cap C = \emptyset$.
\begin{example}
	$(\ar) \ileadstog \phi \lor  (\ar, \ar) \ileadstog \phi \lor …$ is not falsifiable.
	Given any $O$, define $l = \max_{\hist \suchthat \exists p \in P \suchthat (i, \hist, p) \in O} \card{\hist} + 1$. Pick any $\allleadstog$ such that $O \subset {\allleadstog} \land (\ar)^l \ileadstog \phi$. We have ${\allleadstog} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	\label{ex:unfals}
	$(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi$ is not falsifiable.
	Consider any ${\allleadstog} \notin C$, thus satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
	Consider any $O \subseteq {\allleadstog}$ and let us prove that $S_O \cap C ≠ \emptyset$.
	By definition of a general observation, $O$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $O$, pick any ${\allleadstog'}$ such that $O \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have ${\allleadstog'} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	$\hist \ileadstog[j] \phi \land [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
	Consider any $\allleadstog$ satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$, thus with ${\allleadstog} \notin C$.
	\Cref{ex:unfals} concludes.
\end{example}
\begin{example}
	Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
	Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $O = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}

@@

@@

\begin{example}[To be corrected]
Now consider $\allleadsto$ such that $\exists J, K \subset I \suchthat \forall j \in J: \ar \ileadstost[j] \phi \land \forall k \in K: \ar' \ileadstost[k] \phi$.
Then some suitable $C$ exists, for example, $\ar \ileadstost[j] \phi$.
\end{example}

>>>>

We now say that $C$ is \emph{falsifiable} iff $\forall {\allleadstog} \in P^{I × \allhist}: ¬C({\allleadstog}) \implies \exists O \subseteq {\allleadstog} \suchthat O \implies ¬C$, meaning that
$\forall {\allleadstog} \notin C: \exists O \subseteq {\allleadstog} \suchthat S_O \cap C = \emptyset$.
\begin{example}
	$(\ar) \ileadstog \phi \lor  (\ar, \ar) \ileadstog \phi \lor …$ is not falsifiable.
	Given any $O$, define $l = \max_{\hist \suchthat \exists p \in P \suchthat (i, \hist, p) \in O} \card{\hist} + 1$. Pick any $\allleadstog$ such that $O \subset {\allleadstog} \land (\ar)^l \ileadstog \phi$. We have ${\allleadstog} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	\label{ex:unfals}
	$(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi$ is not falsifiable.
	Consider any ${\allleadstog} \notin C$, thus satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
	Consider any $O \subseteq {\allleadstog}$ and let us prove that $S_O \cap C ≠ \emptyset$.
	By definition of a general observation, $O$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $O$, pick any ${\allleadstog'}$ such that $O \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have ${\allleadstog'} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	$\hist \ileadstog[j] \phi \land [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
	Consider any $\allleadstog$ satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$, thus with ${\allleadstog} \notin C$.
	\Cref{ex:unfals} concludes.
\end{example}
\begin{example}
	Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
	Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $O = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}
\begin{example}
	Given an infinite $I$, the claim $\exists i \suchthat \ar \ileadstostg \phi$ is not falsifiable.
	Given any ${\allleadstog} \notin C$, thus such that $\forall i: \ar \nileadstostg \phi$, any $O \subseteq {\allleadstog}$ only has finitely many $(i, \hist, p ≠ \phi)$ thus $S_O \cap C ≠ \emptyset$.
\end{example}

@@

\NewDocumentCommand{\histk}{}{α^{(k)}}
\NewDocumentCommand{\histkend}{}{α^{(k)}_\mathit{end}}
\NewDocumentCommand{\histl}{}{α^{(l)}}
\NewDocumentCommand{\histlend}{}{α^{(l)}_\mathit{end}}
\begin{theorem}[To be checked and proven!]
	Given any $\phi \in P, \ar_k, \ar_l \in \allargs$, consider the following claim $C$: 
	\begin{equation}
		\left\{\begin{aligned}
			&\forall i \in I, \hist \in \allhist \suchthat \ar_k \in \histend \land \hist \nileadstog \phi: \set{\histp \suchthat \hist \subseteq \histp \land \ar_l \in \histpend} \ileadstog \phi;\\
			&\forall i \in I, \hist \in \allhist \suchthat \ar_l \in \histend \land \hist \nileadstog \phi: \set{\histp \suchthat \hist \subseteq \histp \land \ar_k \in \histpend} \ileadstog \phi.
		\end{aligned}\right.
	\end{equation}
	Assume that ${\allleadsto} \in C$. Then 
	$[\ar_k \nileadstost \phi \land \ar_l \nileadstost \phi]$ 
	is equivalent to: 
	\begin{equation}
		\left\{\begin{aligned}
			&\exists \histk \suchthat \ar_k \in \histkend \land \histk \nileadsto \phi \land \set{\histp \suchthat \histk \subseteq \hist' \land \ar_l \in \histpend} \ileadsto \phi;\\
			&\exists \histl \suchthat \ar_l \in \histlend \land \histl \nileadsto \phi \land \set{\histp \suchthat \histl \subseteq \hist' \land \ar_k \in \histpend} \ileadsto \phi.
		\end{aligned}\right.
	\end{equation}
\end{theorem}
\begin{remark}
	Idea: then use contrapositive. Not cons implies some dec arg.
	Prove equivalence to establish that we need to assume not cons to make such claims informative (in such a way).

	As an example, with the shortest possible paths, we get $\histk = (\ar_k)$ and $\histl = (\ar_l)$, thus:
	\begin{equation}
		\left\{\begin{aligned}
			&(\ar_k ) \nileadsto \phi \land \set{\histp \suchthat (\ar_k) \subseteq \histp \land \ar_l \in \histpend} \ileadsto \phi;\\
			&(\ar_l) \nileadsto \phi \land \set{\histp \suchthat (\ar_l) \subseteq \histp \land \ar_k \in \histpend} \ileadsto \phi.
		\end{aligned}\right.
	\end{equation}

	Idea: first consider that $a_l$ equals $a_j \cup a_k$ so that we can reasonably use an axiom.
	Then in the more general case, assume $\phi$ is not in DJ at all, which also implies that $a = a_k \cup a_l$ fails; and we fall back to the same.
\end{remark}
\begin{remark}
	The theorem is realistic in the following sense.
	Some need $a_k$ only; others need $a_j \cup a_k$ (ignorant of $a_j$ at start).
	Consider $C$ saying that if $a_k$ not convincing, then $a_j \cup a_k$ is convincing afterwards and if $a_j \cup a_k$ not convincing, then $a_k$ convincing afterwards. 
	Define $a_l = a_j \cup a_k$.

	In other words.
	Assume that knowing both $a_j$ and $a_k$ is required to get to $\phi$.
	Assume that $J$ knows $a_j$ and when told it, gets bored, thus not convincing. For $j$, only $a_k$ is decisive.
	But $K$ ignores $a_j$. For $k$, only $a_j \cup a_k$ is decisive.
%	
%	Could have:
%	a_k then a_l OK
%	ak KO
%	al KO
%	al then ak OK
\end{remark}

@@

\begin{example}[To be completed]
	Assume that $\exists \ar_j ≠ \ar_k, J, K \subset I \suchthat J \cup K = I \land \card{J} \notin \N \land \card{K} \notin \N \land \forall j \in J: \ar_j \ileadstost[j] \phi \land \forall k \in K: \ar_k \ileadstost[k] \phi \land \forall \ar \in \allargs \setminus \set{\ar_j, \ar_k}, i \in I: \ar \nileadstost \phi$.
	Then no suitable $C$ exists?

	Let’s find some suitable $C$ that shows that $\phi \in P_i$.
\end{example}

Intuition: if $a_j$ and afterwards $a_k$ is convincing, then either $a_j$ is unuseful in which case $a_k$ is decisive; or $a_j$ is useful in which case $a_j \cup a_k$ is decisive. However, could also be that $a_j$ is essential as first argument; thus $a_j$ is useful but $a_j \cup a_k$ is not decisive as $a'$ then $a_j \cup a_k$ is not convincing. Note that the latter part can be excluded by a falsifiable claim: $\nexists a' \suchthat a', a_j \cup a_k \nileadsto \phi$.

The following antecedent excludes two of these three hypothesis, leaving only the one that $a_j$ is useful.
$((a_j), …, a_k)$ convincing ; $a_k$ not decisive (or $(a_k)$ not convincing?); $\nexists \ar \suchthat [(\ar, …, a_k)$ not convincing $\land(\ar, …,  a_j \cup a_k)$ not convincing].
The conclusion is: $a_j \cup a_k$ decisive.

Is this an axiom or is this provable?
And: is the part with $\nexists \ar$ a falsifiable claim?
